---
title: "Pain_Analysis"
author: "Team SpaceX"
date: "12/14/2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r}
library(dplyr)
library(ggplot2)
library(reshape2)
library(readr)
library(caret)
library(randomForest)
library(glmnet)
library(nnet)
library(MASS)
```


## DATA EXPLORATION AND CLEANSING
```{r}
##here::here("SurgicalRegistry.csv")
raw <- read.csv("SurgicalRegistry.csv")
raw[raw==999 | raw == 777] <- NA
# change columns to factors
colsFactors <- c("PatientGender", "PatientRace", "Genetics", "WhatSurgery","Surgeon")
raw[colsFactors] <- lapply(raw[colsFactors], as.factor) 
#Remove patients who haven't had surgery
raw <- raw[-c(12,14,43,46),]
#change scale of predictive scores to match reported scores (out of 10)
raw$PatientPredictive <- raw$PatientPredictive/10
raw$ParentPredictive <- raw$ParentPredictive/10
```


#create analysis data set
```{r}
pain = raw[,c("PatientID","Genetics","PatientGender","Surgeon","PatientAge","PatientWeight","PatientRace","WhatSurgery","SurgeryHX","DODPostOpDay","PatientPredictive","ParentPredictive","Hrs0_12MeanPainScore","Hrs13_24MeanPainScore","Hrs25_36MeanPainScore","Hrs37_48MeanPainScore","Hrs49_60MeanPainScore","Hrs61_72MeanPainScore","Hrs73_84MeanPainScore","Hrs85_96MeanPainScore","Hrs0_12MaxPainScore","Hrs13_24MaxPainScore","Hrs25_36MaxPainScore","Hrs37_48MaxPainScore","Hrs49_60MaxPainScore","Hrs61_72MaxPainScore","Hrs73_84MaxPainScore","Hrs85_96MaxPainScore","Hrs0_12TMorphEquiv","Hrs13_24TMorphEquiv","Hrs25_36TMorphEquiv","Hrs37_48TMorphEquiv","Hrs49_60TMorphEquiv","Hrs61_72TMorphEquiv","Hrs73_84TMorphEquiv","Hrs85_96TMorphEquiv")]
# Replace PatientID 502, 525, 531's self PredicScorest with Median
## because their original value are 808 and 859; PatientID 502 has missing value
pain$PatientPredictive[2] <- median(pain$PatientPredictive[-c(22,28)], na.rm=TRUE)
pain$PatientPredictive[22] <- median(pain$PatientPredictive[-c(22,28)], na.rm=TRUE)
pain$PatientPredictive[28] <- median(pain$PatientPredictive[-c(22,28)], na.rm=TRUE)
# Replace PatientID 502's Parent PredicScore with Median
pain$ParentPredictive[2] <- median(pain$ParentPredictive, na.rm=TRUE)
# Replace missing value in MeanPainScore using MeanPainScore from other time windows
pain[8,]$Hrs13_24MeanPainScore <- mean(c(pain[8,]$Hrs0_12MeanPainScore, pain[8,]$Hrs25_36MeanPainScore))
pain[46,]$Hrs13_24MeanPainScore <- mean(c(pain[46,]$Hrs0_12MeanPainScore,pain[46,]$Hrs25_36MeanPainScore))
pain[31,]$Hrs73_84MeanPainScore <-mean(c(pain[31,]$Hrs61_72MeanPainScore,pain[31,]$Hrs85_96MeanPainScore))
##[16,] Hrs85_96.OfPainScore = 0
pain[16,]$Hrs85_96MeanPainScore <- 0
##[22,]Hrs73_84.OfPainScores = 0
pain[22,]$Hrs85_96MeanPainScore <- 0
pain[22,]$Hrs73_84MeanPainScore <- 0
pain[61,]$Hrs61_72MeanPainScore <- pain[61,]$Hrs49_60MeanPainScore - 1/3 *(pain[61,]$Hrs49_60MeanPainScore - pain[61,]$Hrs85_96MeanPainScore)
pain[61,]$Hrs73_84MeanPainScore <-pain[61,]$Hrs49_60MeanPainScore - 2/3 *(pain[61,]$Hrs49_60MeanPainScore - pain[61,]$Hrs85_96MeanPainScore)
# Replace missing value in MaxPainScore primarily using MaxPainScore from other time windows
pain[8,]$Hrs13_24MaxPainScore <- mean(c(pain[8,]$Hrs0_12MaxPainScore, pain[8,]$Hrs25_36MaxPainScore))
pain[31,]$Hrs73_84MaxPainScore <- mean(c(pain[31,]$Hrs61_72MaxPainScore, pain[31,]$Hrs85_96MaxPainScore))
##[16,] Hrs85_96.OfPainScore = 0
pain[16,]$Hrs85_96MaxPainScore <- 0
##[22,]Hrs73_84.OfPainScores = 0
pain[22,]$Hrs73_84MaxPainScore <- 0
pain[22,]$Hrs85_96MaxPainScore <- 0
## pain[8,]$Hrs49_60MaxPainScore = Hrs85_96MaxPainScore =1
pain[61,]$Hrs61_72MaxPainScore <- 1
pain[61,]$Hrs73_84MaxPainScore <- 1
## pain[24,]$Hrs85_96MinPainScore = Hrs85_96MeanPainScore =3
pain[24,]$Hrs85_96MaxPainScore <- 3
pain$PatientPredictive[pain$PatientPredictive > 10] = pain$PatientPredictive[pain$PatientPredictive > 10]/10
pain$PatientPredictive[pain$ParentPredictive > 10] = pain$PatientPredictive[pain$ParentPredictive > 10]/10
```


#change pain scores to change over previous value

```{r}
# create relative mean pain score change
pain$Hrs0_12MeanPainScore_change <- 0
pain$Hrs13_24MeanPainScore_change <- pain$Hrs13_24MeanPainScore - pain$Hrs0_12MeanPainScore
pain$Hrs25_36MeanPainScore_change <- pain$Hrs25_36MeanPainScore - pain$Hrs13_24MeanPainScore  
pain$Hrs37_48MeanPainScore_change <- pain$Hrs37_48MeanPainScore - pain$Hrs25_36MeanPainScore  
pain$Hrs49_60MeanPainScore_change <- pain$Hrs49_60MeanPainScore - pain$Hrs37_48MeanPainScore  
pain$Hrs61_72MeanPainScore_change <- pain$Hrs61_72MeanPainScore - pain$Hrs49_60MeanPainScore  
pain$Hrs73_84MeanPainScore_change <- pain$Hrs73_84MeanPainScore - pain$Hrs61_72MeanPainScore  
pain$Hrs85_96MeanPainScore_change <- pain$Hrs85_96MeanPainScore - pain$Hrs73_84MeanPainScore  
# create relative max pain score change
pain$Hrs0_12MaxPainScore_change <- 0
pain$Hrs13_24MaxPainScore_change <- pain$Hrs13_24MaxPainScore - pain$Hrs0_12MaxPainScore
pain$Hrs25_36MaxPainScore_change <- pain$Hrs25_36MaxPainScore - pain$Hrs13_24MaxPainScore  
pain$Hrs37_48MaxPainScore_change <- pain$Hrs37_48MaxPainScore - pain$Hrs25_36MaxPainScore  
pain$Hrs49_60MaxPainScore_change <- pain$Hrs49_60MaxPainScore - pain$Hrs37_48MaxPainScore  
pain$Hrs61_72MaxPainScore_change <- pain$Hrs61_72MaxPainScore - pain$Hrs49_60MaxPainScore  
pain$Hrs73_84MaxPainScore_change <- pain$Hrs73_84MaxPainScore - pain$Hrs61_72MaxPainScore  
pain$Hrs85_96MaxPainScore_change <- pain$Hrs85_96MaxPainScore - pain$Hrs73_84MaxPainScore
# create relative morphine change
pain$Hrs0_12TMorphEquiv_change <- 0
pain$Hrs13_24TMorphEquiv_change <- pain$Hrs13_24TMorphEquiv - pain$Hrs0_12TMorphEquiv
pain$Hrs25_36TMorphEquiv_change <- pain$Hrs25_36TMorphEquiv - pain$Hrs13_24TMorphEquiv  
pain$Hrs37_48TMorphEquiv_change <- pain$Hrs37_48TMorphEquiv - pain$Hrs25_36TMorphEquiv  
pain$Hrs49_60TMorphEquiv_change <- pain$Hrs49_60TMorphEquiv - pain$Hrs37_48TMorphEquiv  
pain$Hrs61_72TMorphEquiv_change <- pain$Hrs61_72TMorphEquiv - pain$Hrs49_60TMorphEquiv  
pain$Hrs73_84TMorphEquiv_change <- pain$Hrs73_84TMorphEquiv - pain$Hrs61_72TMorphEquiv  
pain$Hrs85_96TMorphEquiv_change <- pain$Hrs85_96TMorphEquiv - pain$Hrs73_84TMorphEquiv
```

```{r}
# Average change in pain score per patient
pain_change_times1 <- c( "Hrs13_24MeanPainScore_change","Hrs25_36MeanPainScore_change","Hrs37_48MeanPainScore_change","Hrs49_60MeanPainScore_change", "Hrs61_72MeanPainScore_change","Hrs73_84MeanPainScore_change","Hrs85_96MeanPainScore_change")
pain$mean_change_MeanPainScore <- apply(pain[pain_change_times1], 1, mean, na.rm=TRUE)
# ABS change in pain score per patient
pain$mean_absChange_MeanPainScore <- apply(pain[pain_change_times1], 1, function(x) mean(abs(x),na.rm = T))
```

# create vector of time windows to be used in plots
```{r}
times <- c("0-12","13-24","25-36","37-48","49-60","61-72","73-84","85-96")
pain_change_times <- c("Hrs0_12MeanPainScore_change", "Hrs13_24MeanPainScore_change","Hrs25_36MeanPainScore_change","Hrs37_48MeanPainScore_change","Hrs49_60MeanPainScore_change", "Hrs61_72MeanPainScore_change","Hrs73_84MeanPainScore_change","Hrs85_96MeanPainScore_change")
yl <- "Mean Pain Score Change Over Previous 12 Hours"
xl <- "Time Window (Hrs)"
```

# Mean pain times for each 12 hour window
```{r}
mean_pain_time <-colMeans(pain[pain_change_times],na.rm = T)
plot(mean_pain_time,ylab = yl, xlab = xl, xaxt='n',pch=19)
lines(mean_pain_time,col='red')
axis(side=1,at = 1:8, labels=times)
## the largest average difference in pain score over the previous period is in hours 49-60, as there is a relatively large spike in pain difference over the previous 12 hours
```

# Mean pain times for each 12 hour window: Gender
```{r}
mean_pain_time_M <-colMeans(pain[pain$PatientGender==1,pain_change_times],na.rm = T)
mean_pain_time_F <-colMeans(pain[pain$PatientGender==2,pain_change_times],na.rm = T)
plot(mean_pain_time_M, col='blue', type='l', ylab = yl, xlab = xl, xaxt='n',ylim = c(-.3,.5) )
lines(mean_pain_time_F, col='magenta')
axis(side=1,at = 1:8, labels=times)
legend(1, .4, legend=c("Male", "Female"),
       fill=c("blue", "magenta"),cex=0.8)
## Males and females follow roughly teh same trend in terms of reported difference in pain score over time until after the 60th hour post surgery, when males show a considerable drop in pain scores and females have roughly the same pain over the next 24 hours.  
```

# Mean pain times for each 12 hour window: Race
```{r}
mean_pain_time_Wh <-colMeans(pain[pain$PatientRace==1,pain_change_times],na.rm = T)
mean_pain_time_Oth <-colMeans(pain[pain$PatientRace!=1,pain_change_times],na.rm = T)
plot(mean_pain_time_Wh, col='black', type='l',ylim = c(-1,1), ylab=yl, xlab=xl, xaxt='n')
lines(mean_pain_time_Oth, col='red')
axis(side=1,at = 1:8, labels=times)
legend(1, 1, legend=c("White", "Other"),fill=c("black", "red"),cex=0.8)
# Patients with White race report a more steady change in pain scores over the previous period, whereas non white patients have more variance in the reported pain scores compared to the previous 12 hour window
 
```

# Mean pain times for each 12 hour window: Surgeon
```{r}
mean_pain_time_Sg1 <-colMeans(pain[pain$Surgeon==1,pain_change_times],na.rm = T)
mean_pain_time_Sg2 <-colMeans(pain[pain$Surgeon==2,pain_change_times],na.rm = T)
mean_pain_time_SgOth <-colMeans(pain[pain$Surgeon!=2 & pain$Surgeon!=1 ,pain_change_times],na.rm = T)
plot(mean_pain_time_Sg1, col='black', type='l',ylim = c(-1,1), ylab=yl, xlab=xl, xaxt='n')
lines(mean_pain_time_Sg2, col='red')
lines(mean_pain_time_SgOth, col='green')
axis(side=1,at = 1:8, labels=times)
legend(1, 1, legend=c("Thomson", "Lee","Other Surgeon"),fill=c("black", "red","green"),cex=0.8)
## those who were operated on by Thomson show a large spike in pain after 2 days (48 hours).  Those operated by Lee show an overall flatter trend compared to Thomson, but the pain difference over the previous period seems t be higher after 3 days (72 hr)
```

# Mean pain times for each 12 hour window: Surgery
```{r}
mean_pain_time_Nuss <-colMeans(pain[pain$WhatSurgery==1,pain_change_times],na.rm = T)
mean_pain_time_Spine <-colMeans(pain[pain$WhatSurgery==2,pain_change_times],na.rm = T)
plot(mean_pain_time_Nuss, col='black', type='l',ylim = c(-1,1),ylab=yl, xlab=xl, xaxt='n')
lines(mean_pain_time_Spine, col='red')
axis(side=1,at = 1:8, labels=times)
legend(1, 1, legend=c("Chest","Spine"),fill=c("black", "red"),cex=0.8)
# not a large difference in pain difference between the spine and chest surgery patients, as each surgery type experiences higher pain differences over different time windows
```
# Mean pain times for each 12 hour window: Surgery History
```{r}
mean_pain_time_Surg <-colMeans(pain[pain$SurgeryHX==1,pain_change_times],na.rm = T)
mean_pain_time_NoSurg <-colMeans(pain[pain$SurgeryHX==0,pain_change_times],na.rm = T)
plot(mean_pain_time_Surg, col='black', type='l',ylim = c(-1,1),ylab=yl, xlab=xl, xaxt='n')
lines(mean_pain_time_NoSurg, col='red')
axis(side=1,at = 1:8, labels=times)
legend(1, 1, legend=c("Surgery History","No Surgery History"),fill=c("black", "red"),cex=0.8)
# patients with surgery history report a significant drop in pain over teh first 24 hours compared to those without surgery history.  However, after 2 full day, the trends between the two groups are roughly the same
```
# correlate morphine dose with pain score overtime
```{r}
## create data frame to show correlation between pain scored difference and morphine dosage difference for each time window
pain_cor_df <-pain[,grepl( "(Hrs[0-9]+_[0-9]+MeanPainScore_change)|(Hrs[0-9]+_[0-9]+TMorphEquiv_change)",names(pain))]
pain_cor_matrix <- cor(pain_cor_df,use="pairwise.complete.obs")
pain_cor_matrix <- pain_cor_matrix[2:8,10:16]
# create dataset for morphone correlation matrix in tableau
write.csv(pain_cor_matrix,file = "morph.csv")
# there is very low correlation between reported difference in pain scores over the previous period and difference in Morphone dosage over teh previous period across all time windows
```

## T-TESTS AND CORRELATION TESTS ##

#Analysis: Hours 0-12

##Gender

```{r}
p1 <- ggplot(pain, aes(PatientGender, Hrs0_12MeanPainScore)) + geom_boxplot()
p1
t.test(Hrs0_12MeanPainScore ~ PatientGender, pain)
```

##Age

```{r}
p2 <- ggplot(pain, aes(PatientAge, Hrs0_12MeanPainScore)) + geom_point() + geom_smooth()
p2
cor.test(pain$PatientAge, pain$Hrs0_12MeanPainScore, method = "pearson")
```

##Weight

```{r}
p3 <- ggplot(pain, aes(PatientWeight, Hrs0_12MeanPainScore)) + geom_point() + geom_smooth()
p3
cor.test(pain$PatientWeight, pain$Hrs0_12MeanPainScore, method = "pearson")
```

##Race

```{r}
p4 <- ggplot(pain, aes(PatientRace, Hrs0_12MeanPainScore)) + geom_boxplot()
p4
t4 <- pain %>%
  group_by(PatientRace) %>%
  summarise(Hrs0_12MeanPainScore = mean(Hrs0_12MeanPainScore, na.rm=TRUE), count=n())
print(t4)
```

##Genetic Testing

```{r}
p5 <- ggplot(pain, aes(Genetics, Hrs0_12MeanPainScore)) + geom_boxplot()
p5
t.test(Hrs0_12MeanPainScore ~ Genetics, pain)
```

##What Surgery

```{r}
p6 <- ggplot(pain, aes(WhatSurgery, Hrs0_12MeanPainScore)) + geom_boxplot()
p6
t.test(Hrs0_12MeanPainScore ~ WhatSurgery, pain)
```

##SurgeryHX

```{r}
p7 <- ggplot(pain, aes(SurgeryHX, Hrs0_12MeanPainScore)) + geom_boxplot()
p7
t.test(Hrs0_12MeanPainScore ~ SurgeryHX, pain)
```

##PatientPredictive

```{r}
p9 <- ggplot(pain, aes(PatientPredictive, Hrs0_12MeanPainScore)) + geom_point() + geom_smooth()
p9
cor.test(pain$PatientPredictive, pain$Hrs0_12MeanPainScore, method = "pearson")
```

##ParentPredictive

```{r}
p10 <- ggplot(pain, aes(ParentPredictive, Hrs0_12MeanPainScore)) + geom_point() + geom_smooth()
p10
cor.test(pain$ParentPredictive, pain$Hrs0_12MeanPainScore, method = "pearson")
```




#Analysis: Mean Change

##Gender

```{r}
p1 <- ggplot(pain, aes(PatientGender, mean_change_MeanPainScore)) + geom_boxplot()
p1
t.test(mean_change_MeanPainScore ~ PatientGender, pain)
```

##Age

```{r}
p2 <- ggplot(pain, aes(PatientAge, mean_change_MeanPainScore)) + geom_point() + geom_smooth()
p2
cor.test(pain$PatientAge, pain$mean_change_MeanPainScore, method = "pearson")
```

##Weight

```{r}
p3 <- ggplot(pain, aes(PatientWeight, mean_change_MeanPainScore)) + geom_point() + geom_smooth()
p3
cor.test(pain$PatientWeight, pain$mean_change_MeanPainScore, method = "pearson")
```

##Race

```{r}
p4 <- ggplot(pain, aes(PatientRace, mean_change_MeanPainScore)) + geom_boxplot()
p4
t4 <- pain %>%
  group_by(PatientRace) %>%
  summarise(mean_change_MeanPainScore = mean(mean_change_MeanPainScore, na.rm=TRUE), count=n())
print(t4)
```

##Genetic Testing

```{r}
p5 <- ggplot(pain, aes(Genetics, mean_change_MeanPainScore)) + geom_boxplot()
p5
t.test(mean_change_MeanPainScore ~ Genetics, pain)
```

##What Surgery

```{r}
p6 <- ggplot(pain, aes(WhatSurgery, mean_change_MeanPainScore)) + geom_boxplot()
p6
t.test(mean_change_MeanPainScore ~ WhatSurgery, pain)
```

##SurgeryHX

```{r}
p7 <- ggplot(pain, aes(SurgeryHX, mean_change_MeanPainScore)) + geom_boxplot()
p7
t.test(mean_change_MeanPainScore ~ SurgeryHX, pain)
```

##PatientPredictive

```{r}
p9 <- ggplot(pain, aes(PatientPredictive, mean_change_MeanPainScore)) + geom_point() + geom_smooth()
p9
cor.test(pain$PatientPredictive, pain$mean_change_MeanPainScore, method = "pearson")
```

##ParentPredictive

```{r}
p10 <- ggplot(pain, aes(ParentPredictive, mean_change_MeanPainScore)) + geom_point() + geom_smooth()
p10
cor.test(pain$ParentPredictive, pain$mean_change_MeanPainScore, method = "pearson")
```



#Analysis: Absolute Value Change

##Gender

```{r}
p1 <- ggplot(pain, aes(PatientGender, mean_absChange_MeanPainScore)) + geom_boxplot()
p1
t.test(mean_absChange_MeanPainScore ~ PatientGender, pain)
```

##Age

```{r}
p2 <- ggplot(pain, aes(PatientAge, mean_absChange_MeanPainScore)) + geom_point() + geom_smooth()
p2
cor.test(pain$PatientAge, pain$mean_absChange_MeanPainScore, method = "pearson")
```

##Weight

```{r}
p3 <- ggplot(pain, aes(PatientWeight, mean_absChange_MeanPainScore)) + geom_point() + geom_smooth()
p3
cor.test(pain$PatientWeight, pain$mean_absChange_MeanPainScore, method = "pearson")
```

##Race

```{r}
p4 <- ggplot(pain, aes(PatientRace, mean_absChange_MeanPainScore)) + geom_boxplot()
p4
t4 <- pain %>%
  group_by(PatientRace) %>%
  summarise(mean_absChange_MeanPainScore = mean(mean_absChange_MeanPainScore, na.rm=TRUE), count=n())
print(t4)
```

##Genetic Testing

```{r}
p5 <- ggplot(pain, aes(Genetics, mean_absChange_MeanPainScore)) + geom_boxplot()
p5
t.test(mean_absChange_MeanPainScore ~ Genetics, pain)
```

##What Surgery

```{r}
p6 <- ggplot(pain, aes(WhatSurgery, mean_absChange_MeanPainScore)) + geom_boxplot()
p6
t.test(mean_absChange_MeanPainScore ~ WhatSurgery, pain)
```

Closest. p = 0.0815

##SurgeryHX

```{r}
p7 <- ggplot(pain, aes(SurgeryHX, mean_absChange_MeanPainScore)) + geom_boxplot()
p7
t.test(mean_absChange_MeanPainScore ~ SurgeryHX, pain)
```

##PatientPredictive

```{r}
p9 <- ggplot(pain, aes(PatientPredictive, mean_absChange_MeanPainScore)) + geom_point() + geom_smooth()
p9
cor.test(pain$PatientPredictive, pain$mean_absChange_MeanPainScore, method = "pearson")
```

##ParentPredictive

```{r}
p10 <- ggplot(pain, aes(ParentPredictive, mean_absChange_MeanPainScore)) + geom_point() + geom_smooth()
p10
cor.test(pain$ParentPredictive, pain$mean_absChange_MeanPainScore, method = "pearson")
```


## MODELS ##

##  K Means Cluster
```{r}
# repalce NA with median
var_NA = c(names(pain)[37:44])
for(i in var_NA){
  pain[i][is.na(pain[i])] = median(as.numeric(unlist(pain[i])), na.rm = T)
}
p_diff = pain[,c("PatientGender","PatientAge","PatientWeight","WhatSurgery","PatientRace","Surgeon","PatientPredictive","ParentPredictive",names(pain)[37:44])]
k_diff <- kmeans(p_diff, centers = 3)
k_diff
```
```{r}
data = t(matrix(k_diff$centers[,9:15], byrow=F, nrow=3))
matplot(data, type = "l", col = c(1,2,4), xlab = "Hours", ylab = "Pain Diff")
legend("topleft", legend = 1:3, col=c(1,2,4), pch=1)
```
## t-tests among clusters
```{r}
c1 = pain[k_diff$cluster == 1,]
c2 = pain[k_diff$cluster == 2,]
c3 = pain[k_diff$cluster == 3,]
t.test(c1$PatientWeight, c2$PatientWeight)
t.test(c1$PatientWeight, c3$PatientWeight)
t.test(c2$PatientWeight, c3$PatientWeight)
t.test(c1$PatientAge, c2$PatientAge)
t.test(c1$PatientAge, c3$PatientAge)
t.test(c2$PatientAge, c3$PatientAge)
t.test(c1$PatientPredictive, c2$PatientPredictive)
t.test(c1$PatientPredictive, c3$PatientPredictive)
t.test(c2$PatientPredictive, c3$PatientPredictive)
t.test(c1$ParentPredictive, c2$ParentPredictive)
t.test(c1$ParentPredictive, c3$ParentPredictive)
t.test(c2$ParentPredictive, c3$ParentPredictive)
```


## KNN, Elastic Net and Random Forest
```{r}
set.seed(2018)
```

```{r}
pain_idx = sample(1:nrow(pain), size = round(0.7 * nrow(pain)))
pain_train = pain[pain_idx, ]
pain_test = pain[-pain_idx, ]
```

```{r}
# define functions for later use
#get best result
get_best_result = function(caret_fit) {
  best = which(rownames(caret_fit$results) == rownames(caret_fit$bestTune))
  best_result = caret_fit$results[best, ]
  rownames(best_result) = NULL
  best_result
}
#set up accuracy functions
accuracy = function(actual, predicted){
  mean(actual - sd(actual) < predicted & predicted < actual + sd(actual))
}
#use 5-fold cross validation
cv_5 = trainControl(method = "cv", number = 5)
```

```{r}
#get best result
get_best_result = function(caret_fit) {
  best = which(rownames(caret_fit$results) == rownames(caret_fit$bestTune))
  best_result = caret_fit$results[best, ]
  rownames(best_result) = NULL
  best_result
}
```

### K-Nearest Neighbors Model

```{r}
#knn
set.seed(400)
pain_knn = train(
  #try full model
  mean_absChange_MeanPainScore ~ PatientGender + PatientRace + Genetics + WhatSurgery
  + PatientAge + PatientWeight,
  data = pain_train,
  method = "knn",
  trControl = cv_5,
  #choose 34 numbers to be the number of neighbors for KNN model, and it is the largest number we can pick
  tuneGrid = expand.grid(k = seq(1, 25, by = 1))
)
```

```{r}
knn_best = get_best_result(pain_knn)
knn_best
```

```{r}
knn_best[1, 2]
```


```{r}
ncluster = nrow(pain) / knn_best$k
ncluster
```

```{r}
#test accuracy
knn_acc = accuracy(actual = pain_test$mean_absChange_MeanPainScore,
                   predicted = predict(pain_knn, pain_test))
knn_acc
```

### Elastic Net Model

Instead of using logistic regression without regularization, the elastic net model is considered to reduce risk of overfitting. 

```{r}
#fitting linear model, a regularized method that linearly combines penalties of lasso and ridge
pain_elastic = train(
  mean_absChange_MeanPainScore ~ PatientGender + PatientRace + Genetics + 
    WhatSurgery + PatientAge + PatientWeight,
  data = pain_train,
  method = "glmnet",
  trControl = cv_5,
  tuneLength = 6 #largest we can choose
  #if larger, if one of the outcome classes has 0 samples, so sensitivity or specificity is underfined and thus NA
)
```

```{r}
elastic_best = get_best_result(pain_elastic); elastic_best
```

```{r}
elastic_acc = accuracy(actual = pain_test$mean_absChange_MeanPainScore,
                       predicted = predict(pain_elastic, pain_test))
elastic_acc
```


### Random Forest 

```{r}
oob = trainControl(method = "oob")
#set.seed(2018)
pain_rf = train(
  mean_absChange_MeanPainScore ~ PatientGender + PatientRace + Genetics + 
    WhatSurgery + PatientAge + PatientWeight,
  data = pain_train,
  method = "rf",
  metric = "RMSE",
  trControl = oob,
  tuneGrid = expand.grid(.mtry = seq(1, 5, 0.2))
)
```

```{r}
rf_best = get_best_result(pain_rf)
rf_best
```


```{r}
rf_acc = accuracy(actual = pain_test$mean_absChange_MeanPainScore,
                  predicted = predict(pain_rf, pain_test))
rf_acc
```

```{r}
# output a table of results for Random Forest model
#maybe not needed
rf = pain_rf$results
knitr::kable(rf)
```
