---
title: "400_Project"
author: "Chuan Du (Sophie)"
date: "11/13/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Load Data

```{r}
#read in data
raw = read.csv('SurgicalRegistry.csv')
```

### Data Cleansing

```{r}
# replace all 999 and 777 with NA
raw[raw == 999 | raw == 777] <- NA

# change columns to factors
colsFactors <- c("PatientGender", "PatientRace", "Genetics", "WhatSurgery", "Surgeon")
raw[colsFactors] <- lapply(raw[colsFactors], as.factor)

#Remove patients who haven't had surgery
raw <- raw[-c(12,14),]

#change scale of predictive scores to match reported scores (out of 10)
raw$PatientPredictive <- raw$PatientPredictive/10
raw$ParentPredictive <- raw$ParentPredictive/10
```

```{r}
#select column names
pain = raw[,c("Genetics","PatientGender","PatientAge","PatientWeight","PatientRace","WhatSurgery","SurgeryHX","DODPostOpDay","PatientPredictive","ParentPredictive","Hrs0_12MeanPainScore","Hrs13_24MeanPainScore","Hrs25_36MeanPainScore","Hrs37_48MeanPainScore","Hrs49_60MeanPainScore","Hrs61_72MeanPainScore","Hrs73_84MeanPainScore","Hrs85_96MeanPainScore","Hrs0_12MaxPainScore","Hrs13_24MaxPainScore","Hrs25_36MaxPainScore","Hrs37_48MaxPainScore","Hrs49_60MaxPainScore","Hrs61_72MaxPainScore","Hrs73_84MaxPainScore","Hrs85_96MaxPainScore","Hrs0_12TMorphEquiv","Hrs13_24TMorphEquiv","Hrs25_36TMorphEquiv","Hrs37_48TMorphEquiv","Hrs49_60TMorphEquiv","Hrs61_72TMorphEquiv","Hrs73_84TMorphEquiv","Hrs85_96TMorphEquiv")]
```

```{r}
# Replace PatientID 502, 525, 531's self PredicScorest with Median
## because their original value are 808 and 859; PatientID 502 has missing value
pain$PatientPredictive[2] <- median(pain$PatientPredictive[-c(22,28)], na.rm=TRUE)
pain$PatientPredictive[22] <- median(pain$PatientPredictive[-c(22,28)], na.rm=TRUE)
pain$PatientPredictive[28] <- median(pain$PatientPredictive[-c(22,28)], na.rm=TRUE)

# Replace PatientID 502's Parent PredicScore with Median
pain$ParentPredictive[2] <- median(pain$ParentPredictive, na.rm=TRUE)
```

```{r}
# Replace missing value in MeanPainScore using MeanPainScore from other time windows
pain[8,]$Hrs13_24MeanPainScore <- mean(c(pain[8,]$Hrs0_12MeanPainScore, pain[8,]$Hrs25_36MeanPainScore))

pain[46,]$Hrs13_24MeanPainScore <- mean(c(pain[46,]$Hrs0_12MeanPainScore,pain[46,]$Hrs25_36MeanPainScore))
pain[31,]$Hrs73_84MeanPainScore <-mean(c(pain[31,]$Hrs61_72MeanPainScore,pain[31,]$Hrs85_96MeanPainScore))

##[16,] Hrs85_96.OfPainScore = 0
pain[16,]$Hrs85_96MeanPainScore <- 0

##[22,]Hrs73_84.OfPainScores = 0
pain[22,]$Hrs85_96MeanPainScore <- 0
pain[22,]$Hrs73_84MeanPainScore <- 0

pain[61,]$Hrs61_72MeanPainScore <- pain[61,]$Hrs49_60MeanPainScore - 1/3 *(pain[61,]$Hrs49_60MeanPainScore - pain[61,]$Hrs85_96MeanPainScore)
pain[61,]$Hrs73_84MeanPainScore <-pain[61,]$Hrs49_60MeanPainScore - 2/3 *(pain[61,]$Hrs49_60MeanPainScore - pain[61,]$Hrs85_96MeanPainScore)

# Replace missing value in MaxPainScore primarily using MaxPainScore from other time windows

pain[8,]$Hrs13_24MaxPainScore <- mean(c(pain[8,]$Hrs0_12MaxPainScore, pain[8,]$Hrs25_36MaxPainScore))

pain[31,]$Hrs73_84MaxPainScore <- mean(c(pain[31,]$Hrs61_72MaxPainScore, pain[31,]$Hrs85_96MaxPainScore))

##[16,] Hrs85_96.OfPainScore = 0
pain[16,]$Hrs85_96MaxPainScore <- 0

##[22,]Hrs73_84.OfPainScores = 0
pain[22,]$Hrs73_84MaxPainScore <- 0
pain[22,]$Hrs85_96MaxPainScore <- 0


## pain[8,]$Hrs49_60MaxPainScore = Hrs85_96MaxPainScore =1
pain[61,]$Hrs61_72MaxPainScore <- 1
pain[61,]$Hrs73_84MaxPainScore <- 1

## pain[24,]$Hrs85_96MinPainScore = Hrs85_96MeanPainScore =3
pain[24,]$Hrs85_96MaxPainScore <- 3

# Remove patient ID 43 46
## empty Pain Score data
pain <- pain [-c(41,44),]

#change pain scores to change over previous value

# create relative mean pain score change
pain$Hrs0_12MeanPainScore_change <- 0
pain$Hrs13_24MeanPainScore_change <- pain$Hrs13_24MeanPainScore - pain$Hrs0_12MeanPainScore
pain$Hrs25_36MeanPainScore_change <- pain$Hrs25_36MeanPainScore - pain$Hrs13_24MeanPainScore  
pain$Hrs37_48MeanPainScore_change <- pain$Hrs37_48MeanPainScore - pain$Hrs25_36MeanPainScore  
pain$Hrs49_60MeanPainScore_change <- pain$Hrs49_60MeanPainScore - pain$Hrs37_48MeanPainScore  
pain$Hrs61_72MeanPainScore_change <- pain$Hrs61_72MeanPainScore - pain$Hrs49_60MeanPainScore  
pain$Hrs73_84MeanPainScore_change <- pain$Hrs73_84MeanPainScore - pain$Hrs61_72MeanPainScore  
pain$Hrs85_96MeanPainScore_change <- pain$Hrs85_96MeanPainScore - pain$Hrs73_84MeanPainScore  

# create relative max pain score change
pain$Hrs0_12MaxPainScore_change <- 0
pain$Hrs13_24MaxPainScore_change <- pain$Hrs13_24MaxPainScore - pain$Hrs0_12MaxPainScore
pain$Hrs25_36MaxPainScore_change <- pain$Hrs25_36MaxPainScore - pain$Hrs13_24MaxPainScore  
pain$Hrs37_48MaxPainScore_change <- pain$Hrs37_48MaxPainScore - pain$Hrs25_36MaxPainScore  
pain$Hrs49_60MaxPainScore_change <- pain$Hrs49_60MaxPainScore - pain$Hrs37_48MaxPainScore  
pain$Hrs61_72MaxPainScore_change <- pain$Hrs61_72MaxPainScore - pain$Hrs49_60MaxPainScore  
pain$Hrs73_84MaxPainScore_change <- pain$Hrs73_84MaxPainScore - pain$Hrs61_72MaxPainScore  
pain$Hrs85_96MaxPainScore_change <- pain$Hrs85_96MaxPainScore - pain$Hrs73_84MaxPainScore

# create relative morphine change
pain$Hrs0_12TMorphEquiv_change <- 0
pain$Hrs13_24TMorphEquiv_change <- pain$Hrs13_24TMorphEquiv - pain$Hrs0_12TMorphEquiv
pain$Hrs25_36TMorphEquiv_change <- pain$Hrs25_36TMorphEquiv - pain$Hrs13_24TMorphEquiv  
pain$Hrs37_48TMorphEquiv_change <- pain$Hrs37_48TMorphEquiv - pain$Hrs25_36TMorphEquiv  
pain$Hrs49_60TMorphEquiv_change <- pain$Hrs49_60TMorphEquiv - pain$Hrs37_48TMorphEquiv  
pain$Hrs61_72TMorphEquiv_change <- pain$Hrs61_72TMorphEquiv - pain$Hrs49_60TMorphEquiv  
pain$Hrs73_84TMorphEquiv_change <- pain$Hrs73_84TMorphEquiv - pain$Hrs61_72TMorphEquiv  
pain$Hrs85_96TMorphEquiv_change <- pain$Hrs85_96TMorphEquiv - pain$Hrs73_84TMorphEquiv

# Average change in pain score per patient
pain_change_times1 <- c( "Hrs13_24MeanPainScore_change","Hrs25_36MeanPainScore_change","Hrs37_48MeanPainScore_change","Hrs49_60MeanPainScore_change", "Hrs61_72MeanPainScore_change","Hrs73_84MeanPainScore_change","Hrs85_96MeanPainScore_change")

pain$mean_change_MeanPainScore <- apply(pain[pain_change_times1], 1, mean, na.rm=TRUE)


# ABS change in pain score per patient
pain$mean_absChange_MeanPainScore <- apply(pain[pain_change_times1], 1, function(x) mean(abs(x),na.rm = T))

```

```{r}
library(readr)
library(ISLR)
library(class)
library(caret)
library(FNN)
library(randomForest)
library(glmnet)
library(e1071)
library(nnet)
library(MASS)
library(survival)
```


```{r}
set.seed(2018)
```

```{r}
pain_idx = sample(1:nrow(pain), size = round(0.7 * nrow(pain)))
pain_train = pain[pain_idx, ]
pain_test = pain[-pain_idx, ]
```

```{r}
# define functions for later use
#get best result
get_best_result = function(caret_fit) {
  best = which(rownames(caret_fit$results) == rownames(caret_fit$bestTune))
  best_result = caret_fit$results[best, ]
  rownames(best_result) = NULL
  best_result
}

#set up accuracy functions
accuracy = function(actual, predicted){
  mean(actual - sd(actual) < predicted & predicted < actual + sd(actual))
}

#use 5-fold cross validation
cv_5 = trainControl(method = "cv", number = 5)
```


```{r}
#get best result
get_best_result = function(caret_fit) {
  best = which(rownames(caret_fit$results) == rownames(caret_fit$bestTune))
  best_result = caret_fit$results[best, ]
  rownames(best_result) = NULL
  best_result
}
```

### K-Nearest Neighbors Model

```{r}
#knn
set.seed(400)
pain_knn = train(
  #try full model
  mean_absChange_MeanPainScore ~ PatientGender + PatientRace + Genetics + WhatSurgery
  + PatientAge + PatientWeight,
  data = pain_train,
  method = "knn",
  trControl = cv_5,
  #choose 34 numbers to be the number of neighbors for KNN model, and it is the largest number we can pick
  tuneGrid = expand.grid(k = seq(1, 30, by = 1))
)

```


```{r}
knn_best = get_best_result(pain_knn)
knn_best
```



```{r}
ncluster = nrow(pain) / knn_best$k
ncluster
```

```{r}
#test accuracy
knn_acc = accuracy(actual = pain_test$mean_absChange_MeanPainScore,
                   predicted = predict(pain_knn, pain_test))
knn_acc
```


### Elastic Net Model

Instead of using logistic regression without regularization, the elastic net model is considered to reduce risk of overfitting. 

```{r}
#fitting linear model, a regularized method that linearly combines penalties of lasso and ridge
pain_elastic = train(
  mean_absChange_MeanPainScore ~ PatientGender + PatientRace + Genetics + 
    WhatSurgery + PatientAge + PatientWeight,
  data = pain_train,
  method = "glmnet",
  trControl = cv_5,
  tuneLength = 6 #largest we can choose
  #if larger, if one of the outcome classes has 0 samples, so sensitivity or specificity is underfined and thus NA
)
```

```{r}
elastic_best = get_best_result(pain_elastic); elastic_best
```

```{r}
elastic_acc = accuracy(actual = pain_test$mean_absChange_MeanPainScore,
                       predicted = predict(pain_elastic, pain_test))
elastic_acc
```




## Tree (T)

### Random Forest 

```{r}
#random forest
library(randomForest)
```

```{r}
oob = trainControl(method = "oob")
#set.seed(2018)
pain_rf = train(
  mean_absChange_MeanPainScore ~ PatientGender + PatientRace + Genetics + 
    WhatSurgery + PatientAge + PatientWeight,
  data = pain_train,
  method = "rf",
  metric = "RMSE",
  trControl = oob,
  tuneGrid = expand.grid(.mtry = seq(1, 5, 0.2))
)
```

```{r}
rf_best = get_best_result(pain_rf)
rf_best
```


```{r}
rf_acc = accuracy(actual = pain_test$mean_absChange_MeanPainScore,
                  predicted = predict(pain_rf, pain_test))
rf_acc
```

```{r}
#maybe not needed
#rf = pain_rf$results
#knitr::kable(rf)
```

|       Method     |Tuning Parameter|     Parameter Value   |      RMSE    |   Test Accuracy   |
|------------------|----------------|-----------------------|-----------------------|-------------------|
|`KNN` |     "k"        |   `r knn_best[1, 1]`  |  `r knn_best[1, 2]`   |    `r knn_acc`    |
|`Elastic Net` |    "alpha"     |`r elastic_best[1, 1]` | `r elastic_best[1, 3]`|   `r elastic_acc` |
|                  |    "lambda"    |`r elastic_best[1, 2]` |                       |                   |
|`Random Forest` |    "mtry"      |   `r rf_best[1, 3]`   |  `r rf_best[1, 1]`    |    `r rf_acc`     |

