---
title: "Lab Assignment 1"
author: "Chuan Du (Sophie)"
date: "10/30/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem 1

```{r}
#read in table and make matrix
markov = read.table("~/Desktop/400_Lab1/markov100.txt", header = FALSE)
M = as.matrix(markov)
```

#### Problem 1 (a)

```{r, warning=FALSE}
library(expm)
```

```{r}
#generate initial vector (1, 0, ... , 0)
a1 = c(1, rep(0, 99))
#n-step transition, p(10)
prob10 = a1 %*% (M %^% 10)
p_15_10 = prob10[1, 5]; p_15_10
```

**Answer: **  The probability of being in State 5 after 10 transitions $P_{1, 5}(10) = 0.045091$.

#### Problem 1 (b)

```{r}
#initial states for State 1, 2, 3 with 1/3 prob
a_123 = c(rep(1/3, 3), rep(0, 97))
#10-step transitions, get P^n
prob123_10 = a_123 %*% (M %^% 10)
#get P^10[i, j]
prob_123_10 = prob123_10[1, 10]; prob_123_10
```

**Answer: ** The probability of being in State 10 after 10 transitions is **0.08268901**.

#### Problem 1 (c)

```{r}
#calculate pi
Q1 = t(M) - diag(100)
Q1[100, ] = c(rep(1, 100))
rhs = c(rep(0, 99), 1)
Pi = solve(Q1) %*% rhs
Pi[1]
```

**Answer: ** The steady state probability of being in State 1 is **0.01256589**.

#### Problem 1 (d)

$$m = (I - B) ^{-1} e$$

```{r}
#submatrix of M obtained by deleting r and c corresponding state 100
B2 = M[1:99, 1:99]
Q2 = diag(99) - B2
e1 = c(rep(1, 99))
m = solve(Q2) %*% e1; m[1]
```

**Answer: ** The mean first passage time from State 1 to State 100 is **254.9395**.


### Problem 2

```{r}
#read in table and make matrix
page = read.table("~/Desktop/400_Lab1/webtraffic.txt", header = TRUE)
P = as.matrix(page) #1000*81
```

#### Problem 2 (a)

The traffic matrix is shown below.

```{r}
sums = colSums(P)
m_99 = matrix(sums, 9, 9)
#note: t_9i = 0 for all i
Traffic = t(m_99); Traffic
```

#### Problem 2 (b)

The transition probability matrix is shown below.

```{r}
Traffic[9, 1] = 1000
P = Traffic / rowSums(Traffic); P
```

#### Problem 2 (c)

```{r}
Q3 = t(P) - diag(9)
Q3[9, ] = c(rep(1, 9))
rhs2 = c(rep(0, 8), 1)
Pi1 = solve(Q3) %*% rhs2; Pi1
```

#### Problem 2 (d)

```{r}
B3 = P[1:8, 1:8]
Q4 = diag(8) - B3
e8 = c(rep(1, 8))
m8 = solve(Q4) %*% e8; m8
```

```{r}
t2 = c(0.1, 2, 3, 5, 5, 3, 3, 2, 0)
t2 %*% Pi1 * m8[1] #stationary avg time
```

**Answer: ** The average time a visitor spend on the website is **12.25727**.

#### Problem 2 (e)

```{r}
Traffic_new = Traffic
Traffic_new[2, 6] = 0.3 * Traffic_new[2, 3]
Traffic_new[2, 3] = 0.7 * Traffic_new[2, 3]
Traffic_new[2, 7] = 0.2 * Traffic_new[2, 4]
Traffic_new[2, 4] = 0.8 * Traffic_new[2, 4]

P_newtraffic = Traffic_new / rowSums(Traffic_new)
Qe = t(P_newtraffic) - diag(9)
Qe[9, ] = c(rep(1, 9))
rhse = c(rep(0, 8), 1)
Pi2 = solve(Qe) %*% rhse; Pi2
```

```{r}
Pi2 - Pi1
```

**Answer: ** By comparing Pi2 to Pi, we could see that the probability of visiting Page 2 and Page 3 decreases, so that the link works.